A Review of Neutron Imaging

### 2020-10-21 neutron-imaging/neutron.tex
Added another section

\section{Imaging of biologically important structures}\label{s13}
Biological targets have an extremely rich mixture of molecules. A knowledge of
the presence of certain trace elements in the targets reveals important 
information about its functioning, health and pathology. Few imaging techniques
can accurately provide information about the elements. The chemical shift 
imaging in magnetic resonance imaging provides information at the molecular
level. It provides little information about the elements present. This is where
neutron imaging comes into the picture. When neutrons undergo an inelastic
scattering with the nuclei of the target, they nuclei get into an excited
state. They get back to their ground state after emitting a photon. We now 
have a detailed knowledge of the nuclear energy levels. By measuring the energy
of the emitted photon we can guess the nucleus, and thus the element, which
might have produced it. This is the principle behind the neutron stimulated
emission computed tomography (NSECT) imaging.

@@

\section{Conclusion}\label{s14}
In this article we reviewed the physics of neutrons from the perspective of its
role in probing materials at various scales of resolution. We started by 
mentioning the physical properties of neutrons, the way they can be produced,
how they can be detected and how a beam of neutrons can be manipulated. The
electrical neutrality of neutrons allows them to penetrate deeply in materials.
They also have a large collision cross-section with protons, making them
potentially impactful for biological targets. As a result, one must be careful
in exposing living matter to a beam of neutrons. We covered methods to measure
dose of neutron radiation and ways to shield neutrons. Optical and electron
microscopes rely heavily on our ability to manipulate light rays and beams of
electrons. The electrical neutrality of neutrons and their weak intrinsic
magnetic moment make it hard to manipulate their beams. Nevertheless, a few
techniques are available and we covered them in the section on optics.

In the second half of the article we reviewed imaging techniques starting with
the simplest 2-d imaging in which we just record the intensity of the 
transmitted beam and interpret it like an X-ray image. We then covered the
tomographic technique borrowed from X-rays and described its superiority over
2-d imaging in getting the distribution of nuclei in targets. Neutrons also 
have a complex refractive index. The imaginary part of the refractive index is
used to differentiate between nuclei that have similar attenuation properties.
The magnetic moment of neutrons permit them to be treated like protons in the
context of nuclear magnetic resonance. Larmor precession based imaging can be 
used to study the magnetic properties of solids at the level of magnetic 
domains. Neutron spin-echo is another technique borrowed from the world of 
nuclear magnetic resonance. It is used to study the dynamic properties of 
biologically important molecules. We closed the review with a section on the
unique ways in which neutron imaging is used to investigate macro-molecules
and detect elements in biological targets.

>>>>

\section{Neutron imaging of materials}\label{s13}
Neutrons are sensitive to lighter elements and sometimes they have a 
significantly different collision cross section for neighboring elements 
building a contrast between the portions in the images where they are present.
Neutrons are the only probes that can distingush between isotopes. Their
electrical neutrality permits them to penetrate deeply and their small magnetic
moment allows them to interact magnetically with the target. All these
properties make neutrons an attractive probe to investigate the static and
dynamic properties of materials \cite{anderson2008neutron}.

The lost foam casting process consists of molten metal replacing the polystyrene
foam in a cast. Neutrons provide a sharp contrast between the hydrogen-rich
polystyrene and the metal. It is possible to take frequent images of the 
casting process to produce a video of how the molten metal displaces the foam
and settles down in the cast to solidify. One can study the effect of foam
thickness on the quality of casting. The video images also suggest a flow
rate that will not trigger turbulence in the molten metal. They provide 
details of the multiphase flow of the molten metal and the foam that no other
imaging technique gives.

The mechanical properties of materials depend on the local stress fields in
them. Many solids have a non-zero local stress built up during its 
solidification process. These stresses exist in absence of any traction or
forces of constraints. Their presence can be detected from the strains 
accompanying them. We mentioned previously that neutron diffraction is a 
valuable tool to examine crystals. They measure the lattice parameters like
the distance between lattice planes. If $d_0$ is the expected spacing between
lattice plans and if $d$ is the observed value then it indicates the presence
of a strain field of magnitude
\[
\varepsilon = \frac{d - d_0}{d_0}.
\]
The strain field is a tensor. A complete description of it requires one to
measure all its components. Of the nine components, only six are independent.
Once measures the strain at certain orientations $(\theta, \phi)$ and uses
them to infer the six independent cartesian components of the strain tensor.
The relationship between strain at an arbitray orientation $(\theta, \phi)$
and the six cartesian components is
\begin{eqnarray}
\varepsilon(\theta, \phi) &=& \varepsilon_{xx}\cos^2\theta\sin^2\phi +
\varepsilon_{xy}\sin(2\theta)\sin^2\phi \nonumber \\
 & & + \varepsilon_{yy}\sin^2\theta\sin^2\phi+\varepsilon_{zz}\cos^2\phi 
\nonumber \\
 & & + \varepsilon_{xz}\cos\theta\sin(2\phi)+
\varepsilon_{yz}\sin\theta\sin(2\phi) \label{s13e1}
\end{eqnarray}
If we have $\varepsilon(\theta,\phi)$ for six directions then we have six
equations in six unknowns from which we can determine the entire strain field.
Once we have the local strain field, we can compute the local stress field 
using Hooke's law.

The technique to measure local strain and stress described in the previous
paragraph works well for materials that are largely homegeneous. In the case
of granular and porous materials, the stress-strain relations are very 
complicated. They depend on the instantaneous forces as well as past history
of loading. Unlike the homogeneous materials porous and granular materials 
have inhomogenous, anisotropic and non-linear stress-strain fields. Traditional
probes like X-rays cannot penetrate deep inside the material to measure the
fields. However neutron beams can easily reach depths that X-rays cannot making
neutrons a valuable tool to measure stresses and strains in granular media.

\section{Imaging of biologically important structures}\label{s14}
Biological targets have an extremely rich mixture of molecules. A knowledge of
the presence of certain trace elements in the targets reveals important 
information about its functioning, health and pathology. Few imaging techniques
can accurately provide information about the elements. The chemical shift 
imaging in magnetic resonance imaging provides information at the molecular
level. It provides little information about the elements present. This is where
neutron imaging comes into the picture. When neutrons undergo an inelastic
scattering with the nuclei of the target, they nuclei get into an excited
state. They get back to their ground state after emitting a photon. We now 
have a detailed knowledge of the nuclear energy levels. By measuring the energy
of the emitted photon we can guess the nucleus, and thus the element, which
might have produced it. This is the principle behind the neutron stimulated
emission computed tomography (NSECT) imaging.

@@

\section{Conclusion}\label{s15}
In this article we reviewed the physics of neutrons from the perspective of its
role in probing materials at various scales of resolution. We started by 
mentioning the physical properties of neutrons, the way they can be produced,
how they can be detected and how a beam of neutrons can be manipulated. The
electrical neutrality of neutrons allows them to penetrate deeply in materials.
They also have a large collision cross-section with protons, making them
potentially impactful for biological targets. As a result, one must be careful
in exposing living matter to a beam of neutrons. We covered methods to measure
dose of neutron radiation and ways to shield neutrons. Optical and electron
microscopes rely heavily on our ability to manipulate light rays and beams of
electrons. The electrical neutrality of neutrons and their weak intrinsic
magnetic moment make it hard to manipulate their beams. Nevertheless, a few
techniques are available and we covered them in the section on optics.

In the second half of the article we reviewed imaging techniques starting with
the simplest 2-d imaging in which we just record the intensity of the 
transmitted beam and interpret it like an X-ray image. We then covered the
tomographic technique borrowed from X-rays and described its superiority over
2-d imaging in getting the distribution of nuclei in targets. Neutrons also 
have a complex refractive index. The imaginary part of the refractive index is
used to differentiate between nuclei that have similar attenuation properties.
The magnetic moment of neutrons permit them to be treated like protons in the
context of nuclear magnetic resonance. Larmor precession based imaging can be 
used to study the magnetic properties of solids at the level of magnetic 
domains. Neutron spin-echo is another technique borrowed from the world of 
nuclear magnetic resonance. It is used to study the dynamic properties of 
biologically important molecules. We then covered the use of neutron probes to 
measure local stress-strain fields in homogeneous and granular materials. 
We closed the review with a section on the
unique ways in which neutron imaging is used to investigate macro-molecules
and detect elements in biological targets.
